---
Proyecto: "Cross-Selling predicting model"
Recurso: "Development Code"
Autor: Juan Delgado
editor_options: 
  chunk_output_type: console
---

0 - Parameters

```{r}

options(scipen=999) # disable scientific notation

```

1 -  Environment setup 1.1. Library load

```{r message=FALSE, warning=FALSE}
paquetes <- c('data.table',
              'dplyr',
              'tidyr',
              'ggplot2',
              'randomForest',
              'ROCR',
              'purrr',
              'smbinning',
              'rpart',
              'rpart.plot')

instalados <- paquetes %in% installed.packages()

if(sum(instalados == FALSE) > 0) {
  install.packages(paquetes[!instalados])
}
lapply(paquetes,require,character.only = TRUE)

```

1.2 - Data load.

```{r}

df <- fread('datos.csv')


```

2 - Exploratory Data Analysis.
2.1 - General Data and Data type.

```{r echo=TRUE}

as.data.frame(sort(names(df)))
str(df)
glimpse(as_tibble(df))



```

We have May and June variables, but due to project design June will be the target and May will be blinded. So we will get rid of all May variables and those from June not target.

Other variables should be factors: NUM_QUEJAS FALLEC, DUD_SUBJ, SEXO, SEGURO_VIDA, SEGURO_HOGAR, SEGURO_RESTO, SEGURO_AUTO --> recode. For now, we save them as to_factors variable

```{r}

#grepl allow search a string inside other string
patron <- '_MAY|_JUN'
may_june_vars <- names(df)[grepl(patron, names(df))]
may_june_vars <- setdiff(may_june_vars,'SALDO_DEPOS_JUN')
# this variable will be the target so 
```

```{r}

to_factors <- c('NUM_QUEJAS', 'FALLEC', 'DUD_SUBJ', 'SEXO', 
                'SEGURO_AUTO_FEB','SEGURO_AUTO_MAR', 
                'SEGURO_AUTO_ABR','SEGURO_HOGAR_FEB','SEGURO_HOGAR_MAR', 
                'SEGURO_HOGAR_ABR','SEGURO_VIDA_FEB','SEGURO_VIDA_MAR', 
                'SEGURO_VIDA_ABR','SEGURO_RESTO_FEB','SEGURO_RESTO_MAR', 
                'SEGURO_RESTO_ABR')

```

2.2 - Data Quality. Basics Statistics

We get a list of the summary of each variable.
```{r}

lapply(df,summary)

```


2.3 - Data Quality. Null Analysis

```{r}

data.frame(colSums(is.na(df)))

```
Nulls might be converted to zero.

2.4 - Data Quality. Zero analysis.


```{r}

contar_ceros <- function(variable) {
    temp <- transmute(df,if_else(variable==0,1,0))
    sum(temp)
}

num_ceros <- sapply(df,contar_ceros)

# num_ceros <- data.frame(VARIABLE=names(num_ceros),CEROS = num_ceros, stringAsFactors = F) 

# as.numeric(num_ceros),stringsAsFactors = F)

num_ceros <- data.frame(VARIABLE=names(num_ceros),CEROS = as.numeric(num_ceros),stringsAsFactors = F)

num_ceros <- num_ceros %>%
  arrange(desc(CEROS)) %>%
  mutate(PORCENTAJE = CEROS / nrow(df) * 100)
num_ceros


```

